An Early Eocene temperature gradient modelled with geological and ecological proxies

### 2022-10-14 manuscript/manuscript.Rmd
clean scripts and result names

To test whether our logistic regression model can adequately describe latitudinal temperature gradients, we applied a simplified version of the model to modern sea surface temperatures from Bio-Oracle [@...]. As these modern, annual sea surface temperatures estimates are associated with little uncertainty, there is no need for a hierarchical model structure, and we substituted temperature ($t$) for $\mu$ in eq. ? and ?. To verify that the modern gradient can be approximated with limited sampling, we resampled modern sea surface temperatures at modern latitudes corresponding the palaeolatitudes of the Eocene samples. This process was repeated 100 times, randomly chosing a longitude for each latitude in each repetition. The parameters of the non-hierarchical temperature model were estimated for each of the 100 samples, and the iterations after burn-in from the posterior of the parameters were pooled to generate the resulting median temperature gradient with 95 % credible intervals.

@@

Differences between Eocene and modern temperatures at a certain latitude were calculated by randomly pairing all iterations of the posterior from the Eocene and modern temperature gradient model, calculating the Eocene and modern temperature using the respective iterations, taking the difference, and then calculating the median (2.5th, 97.5th percentile) from all pairs of iterations.

Global average temperatures were calculated by taking the weighted mean of the median (2.5th, 97.5th percentile) temperature estimates in 1 degree latitudinal bins. The weights were set to the proportion of global surface area in each latitudinal bin, i.e. decreasing with increasing latitude as:
$$weights = sin(\alpha_{1,i}) - sin(\alpha_{2,i}) ,$$
where $\alpha1$ is the upper, and $\alpha2$ is the lower latitudinal boundary of bin $i$.

@@

```{r results block 1 - for in-line code, echo = FALSE}
setwd('..') # move up to parent folder of /manuscript
motemp_global <- readRDS("results/modern/global_mean_modern.rds")
eotemp_global <- readRDS("results/eeco/global_mean_eeco.rds")

>>>>

To test whether our logistic regression model can adequately describe latitudinal temperature gradients, we applied a simplified version of the model to modern sea surface temperatures from Bio-Oracle [@...]. As these modern, annual sea surface temperatures estimates are associated with little uncertainty, there is no need for a hierarchical model structure, and we substituted temperature ($t$) for $\mu$ in eq. ? and ?. To verify that the modern gradient can be approximated with limited sampling, we resampled modern sea surface temperatures at modern latitudes corresponding the palaeolatitudes of the Eocene samples. This process was repeated 100 times, randomly chosing a longitude for each latitude in each repetition. The parameters of the non-hierarchical temperature model were estimated for each of the 100 samples, and the iterations after burn-in from the posterior of the parameters were pooled to generate the resulting median temperature gradient with 95 % credible intervals ($2.5^{th}$ to $97.5^{th}$ percentile).

@@

Differences between Eocene and modern temperatures at a certain latitude were calculated by randomly pairing all iterations of the posterior from the Eocene and modern temperature gradient model, calculating the Eocene and modern temperature using the respective iterations, taking the difference, and then calculating the median ($2.5^{th}$, $97.5^{th}$ percentile) from all pairs of iterations.

Global average temperatures were calculated by taking the weighted mean of the median ($2.5^{th}$, $97.5^{th}$ percentile) temperature estimates in 1 degree latitudinal bins. The weights were set to the proportion of global surface area in each latitudinal bin, i.e. decreasing with increasing latitude as:
$$weights = sin(\alpha_{1,i}) - sin(\alpha_{2,i}) ,$$
where $\alpha1$ is the upper, and $\alpha2$ is the lower latitudinal boundary of bin $i$.

@@

```{r results block - for in-line code, echo = FALSE}
setwd('..') # move up to parent folder of /manuscript

# add gradient results here

motemp_global <- readRDS("results/modern/global_mean_modern.rds")
eotemp_global <- readRDS("results/eeco/global_mean_eeco.rds")
